## title: "Part 1: Simulation using the exponential distribution"
author: "Ramaji" 
;
date: "19-MAR-2016"
---

#*Project for the "Statistical Inference" course*

### Comparing the simulated mean and variance with the theoretical values

Simulate 1000 groups of 40 exponentials with $\lambda = 0.2$,using a fixed seed, and comparing the distribution of the simulated mean
and variance with the theoretical value of $1 / \lambda$:

```{r results='asis'}
nsim <- 1000
nvals <- 40
lambda <- 0.2
set.seed(123)
simdata <- t(replicate(nsim, rexp(nvals, lambda)))
df <- data.frame(Mean=c(mean(rowMeans(simdata)), 1/lambda),
                 Variance=c(mean(apply(simdata, 1, var)), 1/lambda^2))
rownames(df) <- c("Simulated", "Theoretical")
cat("Mean for Simulataneous=",df$Mean[1]," / Theoritical:",df$Mean[2])
cat("Variance for Simultaneous=",df$Variance[1]," / Theoritical:",df$Variance[2])
```

The simulated value and theoretical value are very close. This is as expected by CLT.

### Assessing if the simulated values are approximately normal

Also, according to the CLT, the distribution of the simulated means should tend to normal. To check this let us normalize the vectors and compare it to a $N(0,1)$ distribution.

```{r}
library(ggplot2)
meanvals <- rowMeans(simdata)
zmean <- (meanvals - mean(meanvals)) / sd(meanvals)
qplot(zmean, geom = "blank") +
    geom_line(aes(y = ..density.., colour = 'Empirical'), stat = 'density') +  
    stat_function(fun = dnorm, aes(colour = 'Normal')) +                       
    geom_histogram(aes(y = ..density..), alpha = 0.4, binwidth=.35) +
    geom_vline(xintercept=0, colour="red", linetype="longdash") +
    scale_colour_manual(name = 'Density', values = c('red', 'blue')) +
    ylab("Density") + xlab("z") + ggtitle("Mean values distribution") +
    theme_bw() + theme(legend.position = c(0.85, 0.85))
```

### Evaluating the coverage of the confidence interval

Theoretically, a 95% confidence interval should contain, if we simulate a big number of them, the mean value for the exponential distribution ($1 / \lambda$) 95% of the time.

```{r}
set.seed(123)
lambda <- 0.2
# checks for each simulation if the mean is in the confidence interval
conf <- function(lambda) {   
    ehats <- rexp(1000, lambda)
    se <- sd(ehats)/sqrt(1000)
    ll <- mean(ehats) - 1.96 * se
    ul <- mean(ehats) + 1.96 * se
    (ll < 1/lambda & ul > 1/lambda)   #Returns the limits
}
# estimate the mean in each round of simulations
avgs <- function(lambda) {
    vals <- replicate(100, conf(lambda))
    mean(vals) # Returns the mean
}
# perform the simulation
sim <- replicate(100, avgs(lambda))
mn<-mean(sim)
```
```{r echo=FALSE}
cat("Theoritical CI:",mn)
```
This shows the resulting theoritical confidence interval is close to the expected 95%.
